---
title: "how_to_stargaze_challenge_solution"
output: pdf_document
date: 05-09-22
author: daniel turner (dturner at northwestern at u dot edu)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages
library( stargazer )
library( tidyverse )

# install this package to download the data
# install.packages('fueleconomy')

# load the data
more_cars = fueleconomy::vehicles # the new data is bundled inside ggplot2
og_cars = datasets::mtcars        # if you want to compare to previous dataset

```


# Stargazer challenge question
To get personal experience using `stargazer`, please follow along with this coding challenge. If you get stuck, or want to see my sample solution, check out the file *stargaze_challenge_solution.Rmd*. 

Here's everything you need to know to get started:
1. In Part A, you will create a table from a dataframe for showing raw data
2. In Part B, your task will be to show a stargazer summary statistics table
3. In Part c, you will fit and summarize a statistical model

## About the more_cars data
The data is similar to the Motor Trend car data from the tutorial, but updated with many more cars and variables. Its source is the EPA, collected from 1985-2015, and it excludes vehicles with missing data. The data is about 1000x larger than the *MTcars*.


## Part A: Raw data
Edit the code chunk below so that the result is 2 tables with the head() of both `more_cars` and `og_cars`, which will give you a sense for how these data are similar, and different. If you want to sample N rows instead of taking the first N rows, the code is in the tutorial.


```{r stargazer_challenge_A}

# Make a table of the raw data (only the head)
stargazer( data = more_cars %>% head(),    # let the data be the head of more_cars
           # title = "More cars raw data", # add a title for each table
           type = "text"                   # design in "text" mode, output in whatever format you need
)

# Same for og_cars
stargazer( data = og_cars %>% head(),    # let the data be the head of og_cars
           type = "text"                 
)
```

## Part B: Summary stats
In this code chunk you will make a summary statistics table for `more_cars` with customizations: (1) Omit the column with the "id" field, which only makes sense if joining to other EPA data; (2) remove the "," separating thousands by settings the `digits.separator` argument, and round all summary stats to the nearest whole number by stting the `digits` argument. See the comments for more details.

```{r stargazer_challenge_B}

# Basic recipe for the default summary statistics table
stargazer( data = og_cars,
           type = "text")

# Here are the arguments you will need to customize the output of more_cars:
# omit =                # let 'omit' be 'c("id")'
# digit.separator =     # let 'digit.separator' be ""
# digits =              # let 'digits' be 0 to 

# Result
stargazer( data = more_cars,  
           type = "text"
)


```

## Part C: Model summary
Before we can go about making tables of model summaries, we will need some models to visualize.

In the code chunk below, we will create 2 Generalized Linear Models (GLMs), one predicting highway fuel efficiency and another for city efficiency. The function works a lot like lm() but the output will give access to different summary statistics, because of how the regression is implemented.

```{r stargazer_challenge_C_model}

# Run this chunk to populate the two linear regression models
more_cars_model_hwy = glm( formula = hwy ~ cyl + displ, data = more_cars )
more_cars_model_cty = glm( formula = cty ~ cyl + displ, data = more_cars)

```


Now that we have our models fit to the data, we will visualize them side-by-side, with customizations that make sense for this data. For this table, you will specify the `covariate.labels`, `column.labels`, while hiding the `model.numbers` and dependent variables (`dep.var.labels.include`). See the comments for an explanation of how exactly to set each argument.


```{r stargazer_challenge_C_output}

# Here are the arguments you will need:
# covariate.labels =       # vector of labels to override the covariate labels (cylinders, displacement, intecept/constant)
# column.labels =          # vector of labels to be displayed above our to models, in lieu of model.numbers
# model.numbers =          # let 'model.numbers' be FALSE to hide (1) and (2) over our models
# dep.var.labels.include = # let 'dep.var.lablels.include' be FALSE to hide redundant labels in this case

# Putting all the pieces together to make a table of both model outputs
stargazer( more_cars_model_hwy, more_cars_model_cty,     # The models
           type = "text")       


```

[end]